---
title: "WILD7150 Lecture and Notes"
author: "Steven Cabrera"
date: "Fall 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(dplyr)
```

## Wednesday, August 17, 2022

## Friday, August 19, 2022

## Monday, August 22, 2022

## Wednesday, August 24, 2022

Regression analysis in R.   
Example: y = biomass (kg), x = rainfall (cm)    
True relationship would be: y = 2 +3x + $~N(o,sigma)  
B0 is usually a nuisance parameter, only really useful to best fit the line.  
B1 = slope  
r2, variation in y explained in x  
adjusted r2 takes r2 based on number of x variables in model, in truth x variable is added, r2 goes up. **use multiple r2 when reporting in labs**  
function confint only produces limits. need to do math to find CI.
upper minus lower / 2 **OR** upper - b1, b1- lower, report only one  

Generic sentence  
 
**For each 1 [x-unit] increase in [x], we found a [B1] [y-unit] [+/- 95% CI]; 95% CI [increase/decrease] in [y] ([p-value, r^2^])**  
Include units, specify you gave the CI (some people give SE), for unit rounding - only report two significant figs, use increase/decrease to specify pos or neg slope.  
If necessary - scale data to make data more interpretable.   

```{r, echo = F, message = F, render = F}
data <- read_csv("Lab Data/Lab1/Assign1a.csv")
res <- lm(data = data, WolfSpiders ~ PreySpiders)
summary(res)
anova(res)
plot <- plot(WolfSpiders ~ PreySpiders, data = data)
```


## LAB - Wednesday, August 24, 2022
 
Not much going on. 

## Friday, August 26, 2022

Scaling is very useful for conveying story message/interpreting the *biological signifigance*   
P-value: probability of getting observed results when the null is true.  
Type I Error: detecting a significant effort when there is none (will happen about 5% of the time).  


**Assumptioms of GLM**  
If you ignore/violate assumptions,  
  1. Slopes usually aren't effected  
  2. P-values and CIs are conservative (i.e., they are bigger)  
  3. Linear models are robust to violating of assumptions.  
**Actual assumptions**    
 * 1. Continuous Y  
 * 2. Error is normally distributed - 
 * 3. Linear relationship between X and Y     
    + Need to test or visualize non-linear relationships. Basal area on other things?  
  4. Homoscodastiscity of variance  - unequal variance
  5. No autocorrelation/independent samples - one data point in Y is related to the next or subsequent Y data point



## Monday, August 29, 2022

## Wednesday, August 31, 2022

## LAB - Wednesday, August 31, 2022
