---
title: "Lab 5"
author: "Steven Cabrera"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(readr)
library(ggplot2)
library(dplyr)
library(car)
library(carData)
```

## Dataset #1 

```{r, include = F, echo = F}
data1 <- read_csv("Lab Data/Lab5/Assignment4b.csv")
res1 <- lm(data = data1, SqDens~RoadDensity+Forest+RoadDensity:Forest)
summary(res1)
pine <- filter(data1, Forest=="Pine")
pineres <- lm(data=pine, SqDens~RoadDensity)
summary(pineres)
confint(pineres)
hardwood <- filter(data1, Forest=="Hardwood")
hardwoodres <- lm(data=hardwood, SqDens~RoadDensity)
summary(hardwoodres)
confint(hardwoodres)
ggplot(data1, aes(x=RoadDensity, y = SqDens, color = Forest)) +
  geom_smooth(method = "lm")

```


1. $Y = B_0 + B_1 RoadDensity + B_2 Habitat + B_3 RoadDensity * Habitat + E \sim \text{N}(\mu, \sigma)$  

    * $B$ variables:
        + $B_0$: effect on squirrel density when all x's are 0
        + $B_1$: effect of road density on squirrel density in hardwood
        + $B_2$: difference in squirrel density between pine and hardwood habitat 
        + $B_3$: the effect of road density and differences in pine and hardwood habitat on squirrel density
        
2. We observed a significant interaction between pine and hardwood habitats (p = 3.70e-09), therefore we will analyze the two groups separately. 

3. * In pine habitats, for each 1 road/sq. km increase in road density, we observed a 0.99 squirrels/ha (± 0.13; 95% C.I.) decrease in squirrel density (p = 3.3e-15, r^2^ = 0.89).  
    * In hardwood forests, for each 1 road/sq. km increase in road density, we observed a 0.47 (± 0.13; 95% C.I.) decrease in squirrel density (p = 4.94e-8, r^2^ = 0.66).
    
## Dataset #2
 
```{r, include = F, echo = F}
data2 <- read_csv("Lab Data/Lab5/Assignment4c.csv")
res2 <- lm(data = data2, Size~Sex+Treatment+Sex:Treatment)
summary(res2)
male <- filter(data2, Sex=="Males")
maleres <- lm(data=male, Size~Treatment)
summary(maleres)
confint(maleres)
fem <- filter(data2, Sex=="Females")
femres <- lm(data=fem, Size~Treatment)
summary(femres)
confint(femres)
```

1. $Y = B_0 + B_1 Sex + B_2 Treatment + B_3 Sex * Treatment + E \sim \text{N}(\mu, \sigma)$ 

    * $B$ variables:
        + $B_0$: effect on size when all x's are 0
        + $B_1$: difference in sex on size 
        + $B_2$: difference in treatment on size 
        + $B_3$: treatment difference in size between males and females

2. We observed a significant interaction between sex and treatment (p = 3.06e-09), therefore we will analyze the two groups separately. 

3. * In males, we found that those using hormone treatment had 0.96 (± 0.11; 95% C.I.) increase in size (p = < 2e-16) as opposed to placebo.  
  * In females, we found that those using hormone treatment had 0.47 (± 0.10; 95% C.I.) increase in size (p = 3.25e-12) as opposed to placebo.
    
## Dataset #3
 
```{r, include = F, echo = F}
data3 <- read_csv("Lab Data/Lab5/Assignment4d.csv")
res3 <- lm(data=data3, Size~Elevation+Latitude+Elevation:Latitude)
summary(res3)
confint(res3)
res4 <- lm(data=data3, Size~Elevation)
summary(res4)
confint(res4)
res5 <- lm(data=data3, Size~Latitude)
summary(res5)
confint(res5)
```


1. $Y = B_0 + B_1 Elevation + B_2 Latitude + B_3 Elevation * Latitude + E \sim \text{N}(\mu, \sigma)$ 

2. * $B$ variables:
        + $B_0$: effect on size when all x's are 0
        + $B_1$: effect of elevation on size 
        + $B_2$: effect of latitude on size 
        + $B_3$: effect of elevation and latitude on size
        
3. For each 100 m increase in elevation, we observed a 0.125 (± 0.0567; 95% C.I.) cm decrease in size at zero latitude (p = 3.10e-05). For each 100 degree change in latitude, we observed a 0.17 (± 0.0517; 95% C.I.) cm decrease in size at zero elevation (p = 4.79e-09). We found a significant interaction (p = < 2e-16), such that for each 1 m increase in elevation, we observed that the slope of the latitude-size relationship increased by 0.001005 cm/degrees (± 3.6e-07; 95% C.I.).  