---
title: "WILD7150 Lab 2 - Assumptions of Regression"
author: "Steven Cabrera"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(dplyr)
```

## Dataset #1

```{r, include = F, echo = F}
data1 <- read_csv("Lab Data/Lab2/Ass3Data1.csv")
plot(PredRate~PreyDens, data = data1)
res1 <- lm(PredRate~PreyDens, data = data1)
plot(res1)
summary(res1)
residuals(res1)
plot(residuals(res1)~data1$PreyDens)
confint(res1)
abline(res1)
hist(data1$PredRate)

```

For each 1 prey/ha increase in prey density, we observed a 0.41 prey killer per predator per day (± 0.19; ± 95% C.I.) increase in predation rate (p = 3.23e-05; r^2^ = 0.11). 

**Violates assumption of homoscedasticity**    

```{r, echo = FALSE, message = F}
plot(PredRate~PreyDens, data = data1)
abline(res1)
```

## Dataset #2

```{r, include = F, echo = F}
data2 <- read_csv("Lab Data/Lab2/Ass3Data2.csv")
plot(BasalArea~Age, data = data2)
res2 <- lm(BasalArea~Age, data = data2)
plot(res2)
residuals(res2)
plot(residuals(res2)~data2$Age)
summary(res2)
confint(res2)
abline(res2)
hist(data2$BasalArea)
acf(residuals(res2)) [order(data2$Age)]

```

For each 1 year increase in stand age, we observed a 1.83 m2 (± 0.079; ± 95% C.I.) increase in basal area (p = <2e-16; r^2^ = 0.97). 

**Violates assumption of no autocorrelation**  

```{r, echo = FALSE, message = F}
acf(residuals(res2)) [order(data2$Age)]
```


## Dataset #3

```{r, include = F, echo = F}
data3 <- read_csv("Lab Data/Lab2/Ass3Data3.csv")
plot(PreyDens~PredDens, data = data3)
res3 <- lm(PreyDens~PredDens, data = data3)
plot(res3)
residuals(res3)
plot(residuals(res3)~data3$PredDens)
summary(res3)
confint(res3)
abline(res3)
hist(data3$PreyDens)
acf(residuals(res3)) [order(data3$PredDens)]

```

For each 1 predators/ha increase in predator density, we observed a 1.24 prey/ha (± 0.43; ± 95% C.I.) decrease in prey density (p = 6.13e-07; r^2^ = 0.41). 

**No assumptions violated**  
```{r, echo = FALSE, message = F}
plot(PreyDens~PredDens, data = data3)
abline(res3)
```

## Dataset #4

```{r, include = F, echo = F}
data4 <- read_csv("Lab Data/Lab2/Ass3Data4.csv")
plot(LitterSize~PreyDens, data = data4)
res4 <- lm(LitterSize~PreyDens, data = data4)
plot(res4)
residuals(res4)
plot(residuals(res4)~data4$PreyDens)
abline(a=0,b=0)
summary(res4)
confint(res4)
abline(res4)
hist(data4$LitterSize)
acf(residuals(res4)) [order(data4$PreyDens)]

```

For each 1 prey/ha increase in prey density, we observed a 0.37 pup (± 0.17; ± 95% C.I.) increase in coyote litter size (p = 5.75e-05; r^2^ = 0.15). 

**Violates assumption of normality**  

```{r, echo = FALSE, message = F}
plot(residuals(res4)~data4$PreyDens)
abline(a=0,b=0)
hist(data4$LitterSize)
```

## Dataset #5

```{r, include = F, echo = F}
data5 <- read_csv("Lab Data/Lab2/Ass3Data5.csv")
plot(ProbUse~RoadDens, data = data5)
res5 <- lm(ProbUse~RoadDens, data = data5)
plot(res5)
residuals(res5)
plot(residuals(res5)~data5$RoadDens)
abline(a=0,b=0)
summary(res5)
confint(res5)
abline(res5)
hist(data5$ProbUse)
acf(residuals(res5)) [order(data5$RoadDens)]

```

**Though it is a non-significant relationship**, for each 1 road/km^2^ increase in road density, we observed a 0.0011 percent (± 0.17; ± 95% C.I.) decrease in probability of habitat use by wolves (p = 0.738; r^2^ = 0.0012). 

**Violates assumption of linearity**  

```{r, echo = FALSE, message = F}
plot(ProbUse~RoadDens, data = data5)
abline(res5)
```
