---
title: "Lab 4"
author: "Steven Cabrera"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(readr)
library(ggplot2)
library(dplyr)
library(car)
library(carData)
library(emmeans)
```

## Dataset #1 

```{r, include = F, echo = F}
data1 <- read_csv("Lab Data/Lab4/Assignment4.csv")
res1 <- lm(data = data1, Size~Elevation+Latitude+as.factor(Country))
res2 <- lm(data = data1, Size~Elevation+Latitude)
anova(res1,res2)
summary(aov(res1))
summary(res1)
confint(res1)
data1$Country <- factor(data1$Country, levels = c("Nepal","Tibet","India"))
```
In this example, you have two continuous factors (e.g., elevation [meters] and degrees north latitude [degrees]), and one categorical factor with three groups (country). The dependent variable is lotus plant size (grams). In this system, there is no collinearity among your variables.

1. $Y = B_0 + B_1 X_1 + B_2 X_2 + B_3 X_3 + E \sim \text{N}(\mu, \sigma)$  
    * X varibles:
        + $X_1$ = Latitude (continuous, degrees)
        + $X_2$ = Elevation (continuous, meters)
        + $X_3$ = Country (categorical, three groups = "India", "Nepal", "Tibet" ; reference = "India" in this case)
    
2. $B$ variables:
    * $B_0$: y-intercept
    * $B_1$: slope (estimate) of latitude
    * $B_2$: slope (estimate) of elevation
    * $B_3$: slope (estimate) of country
    
4. The significance of "Country" as a whole is $< 2.2e^{-16}$.  
5. Example sentences:
    * For each 100 m increase in elevation, we observed a 0.162 (± 0.024; 95% C.I.) grams decrease in lotus plant size (p = < 2e-16).
    * For each 0.1 latitude increase, we observed a .1006 (± 0.017; 95% C.I.) grams decrease in lotus plant size (p = < 2e-16).
    * We found that Nepal had 3.59 (± 051; 95% C.I.) g less in lotus plant size than India (p = < 2e-16).
    * We found that Tibet had 1.73 (± 0.49; 95% C.I.) g less in lotus plant size than India (p = < 4.39e-10)
    * We found that Tibet had 1.86 (± 0.50; 95% C.I.) g more in lotus plant size than Nepal (p = 4.92e-11).

## Dataset #2 

```{r, include = F, echo = F}
data2 <- read_csv("Lab Data/Lab4/Assignment4.1.csv")
plot(data=data2, Food~Understory)
res2 <- lm(Food~Understory, data=data2)
summary(res2)
abline(res2)
res2one <- lm(Bunnies~Food, data = data2)
summary(res2one)
confint(res2one)
res2two <- lm(Bunnies~Understory, data = data2)
summary(res2two)
confint(res2two)
bigone <- lm(Bunnies~Understory+Food, data = data2)
summary(bigone)
confint(bigone)
vif(bigone)
```


2.  
```{r, echo = F, message = F}
plot(data=data2, Food~Understory)
abline(res2)
```

3. The $r^2$ between food and understory is 0.7746. 

4.  
```{r, echo = F, message = F}
plot(data=data2, Bunnies~Understory)
```

5.  
```{r, echo = F, message = F}
plot(data=data2, Bunnies~Food)
```

6. For each 1 kg of browse per square meter increase, we observed a 2.78 (± 0.087; 95% C.I.) bunnies/ha increase in density (p = < 2e-16, $r^2$ = 0.976).  

7. For each 0.1 percent understory cover increase, we observed a 0.905 (± 0.065; 95% C.I.) bunnies/ha increase in density (p = <2e-16, $r^2$ = 0.885).

8. For each 0.1 percent understory cover increase, we observed a 0.304 (± 0.014; 95% C.I.) bunnies/ha increase in density (p = <2e-16, $r^2$ = 0.99). For each 1 kg of browse per square meter increase, we observed a 2.00 (± 0.042; 95% C.I.) bunnies/ha increase in density (p = < 2e-16, $r^2$ = 0.99).  

9. The VIF for understory AND food is 4.43751.

10. The coefficient estimates in the multiple regression model capture truth more accurately. The model estimates the effect of understory as 3.04 (truth = 3) as opposed to 9 in a simple regression. The model estimates the effect of understory as 1.997 (truth = 2) as opposed to 2.78 in a simple regression. I would include both variables in a model, while stating that these two variables are potentially confounding. I've used this exercise as a reminder of the utility of plotting data to visualize individual relationships to compare effect sizes, as well as the combined effect of including multiple variables in a model. 

 

## Dataset #3 

```{r, include = F, echo = F}
data3 <- read_csv("Lab Data/Lab4/Assignment4.2.csv")
res5 <- lm(data=data3, Organic~Sediment)
summary(res5)
res6 <- lm(data=data3, Clarity~Sediment)
res7 <- lm(data=data3, Clarity~Organic)
summary(res6)
confint(res6)
summary(res7)
confint(res7)
res8 <- lm(data=data3, Clarity~Sediment+Organic)
summary(res8)
confint(res8)
vif(res8)
```

2.  
```{r, message = F, echo = F}
plot(data = data3, Organic~Sediment)
```

3. The $r^2$ between sediment and organic matter is 0.7289. 

4.  
```{r, echo = F, message = F}
plot(data=data3, Clarity~Sediment)
```

5.  
```{r, echo = F, message = F}
plot(data=data3, Clarity~Organic)
```

6. For each 1 gram/cm3 of sediment, we observed a 0.404 (± 0.016; 95% C.I.) cm decrease in clarity (p = < 2e-16, $r^2$ = 0.9635).  

7. For each 1 gram/cm3 of organic matter, we observed a 0.32 (± 0.041; 95% C.I.) cm decrease in clarity (p = <2e-16, $r^2$ = 0.706).

8. For each 1 gram/cm3 of sediment, we observed a 0.4007 (± 0.0304; 95% C.I.) cm decrease in clarity (p = < 2e-16). For each 1 gram/cm3 of organic matter, we observed a 0.0034 (± 0.028; 95% C.I.) cm decrease in clarity (p = 0.81 **non-significant value**).  

9. The VIF for sediment AND organic is 3.688.  

10. The coefficient estimate of sediment (which is a better "index" of run-off) did not change much when modeled as a single regression compared to multiple regression with organic matter. In both bases, the coefficient is very close to truth (0.4 g/cm3). As the coefficient does not change, the collinear organic matter variable can be considered "redundant" and removed from the model, indicating no observable impact on run-off. I've learned that applying multiple models with the same data can be useful in exploratory analyses for eliminating redundant variables in collinear data.  
 