---
title: "Lab 6"
author: "Steven Cabrera"
date: "2022-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(readr)
library(ggplot2)
library(dplyr)
library(car)
library(carData)
library(nlme)
```

## Dataset #1 

```{r, include = F, echo = F}
data1 <- read_csv("Lab Data/Lab6/AssignA.csv")
res1 <- lm(data = data1, Density~Plot)
summary(res1)
confint(res1)
res1a <- lme(data = data1, Density~Plot, random = ~1|Field)
summary(res1a)
intervals(res1a)
data1$Plot <- factor(data1$Plot, levels = c("dBurn","Control","Fertilize"))
```

1. We found that burned plots had 5.3 (± 3.93; 95% C.I.) hoppers/ha more than control plots (p = 0.0104), while fertilizer plots had 5.20 (± 3.93; 95% C.I.) hoppers/ha more than control plots (p = 0.012). Fertilizer plots had 0.14 (± 3.93; 95% C.I.) hoppers/ha less than burned plots (p = 0.94 **non-significant relationship**.  

2. We found that burned plots had 5.3 (± 3.27; 95% C.I.) hoppers/ha more than control plots (p = 0.0036), while fertilizer plots had 5.20 (± 3.27; 95% C.I.) hoppers/ha more than control plots (p = 0.0044). Fertilizer plots had 0.14 (± 3.27; 95% C.I.) hoppers/ha less than burned plots (p = 0.93) **non-significant relationship**.  

3. The standard deviation due to random effect of Field is 2.23. 

4. The lme model increases power (smaller confidence intervals and lower p-values).  


## Dataset #2 

```{r, include = F, echo = F}
data2 <- read_csv("Lab Data/Lab6/AssignB.csv")
res2 <- lm(data = data2, Size~Fertilizer*Hormone)
summary(res2)
confint(res2)
res2a <- lm(data = data2, Size~Fertilizer+Hormone)
anova(res2,res2a)
summary(res2a)
confint(res2a)
res2b <- lme(data = data2, Size~Fertilizer*Hormone, random = ~1|Field)
summary(res2b)
intervals(res2b)

```


1. We observed a **non-significant** interaction between Fertilizer and Hormone (p = 0.94). Using a fertilizer treatment a 9.03 (± 6.12; 95% C.I.) cm increase in tree size (p = 0.0052), compared to a no fertilizer treatment. Using a hormone treatment had a 2.5 (± 6.12; 95% C.I.) cm increase in tree size (p = 0.408) compared to a no hormone treatment.  

2. We observed a **non-significant** interaction between Fertilizer and Hormone (p = 0.56). Using a fertilizer treatment a 9.03 (± 0.75; 95% C.I.) cm increase in tree size (p = 0.000001), compared to a no fertilizer treatment. Using a hormone treatment had a 2.51 (± 0.75; 95% C.I.) cm increase in tree size (p = 0.000001) compared to a no hormone treatment.

3. The standard deviation due to random effect of Field is 8.55. 

4. The lme model increases power (smaller confidence intervals and lower p-values).  

## Dataset #3
```{r, include = F, echo = F}
data3 <- read_csv("Lab Data/Lab6/AssignC.csv")
res3 <- lm(data = data3, Size~Age)
summary(res3)
confint(res3)
res3a <- lme(data = data3, Size~Age, random = ~1|Individual)
summary(res3a)
intervals(res3a)
res3b <- lme(data = data3, Size~Age, random = ~1|Individual,correlation = corARMA(p = 0, q = 1))
summary(res3b)
intervals(res3b)
anova(res3b,res3a)
```

1. For each 1 year, we observed a 1.02 (± 0.68; 95% C.I.) m increase in tree size (p = 0.0038).  

2. For each 1 year, we observed a 1.02 (± 0.55; 95% C.I.) m increase in tree size (p = 3e-4).  

For each 1 year, we observed a 1.02 (± 0.607; 95% C.I.) m increase in tree size (p = 1e-3) **running average autocorrelated**

3. The standard deviation due to random effect of Field is 9.00.

4. When modeling a moving average autocorrelation, we observed a Theta of 0.13 and standard deviation due to random effects 8.83. An F-drop test comparing models showed a **non-signifcant** improvement to the model (p = 0.16). 

5. The lme model increases power (smaller confidence intervals and lower p-values).  