---
title: "WILD7150 Lab 3 - Categorical X"
author: "Steven Cabrera"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(readr)
library(ggplot2)
library(dplyr)
```

## Dataset #1 

```{r, include = F, echo = F}
data1 <- read_csv("Lab Data/Lab3/Ass5Data1.csv")
boxplot(QuailDensity~Plots, data = data1)
res1 <- lm(QuailDensity~Plots, data = data1)
summary(res1)
confint(res1)
abline(res1)

```

We found that exclusion plots had 3.39 (± 0.59; ± 95% C.I.) quail/ha greater than control plots (p = 1.73e-15, r^2^ = 0.74). 

## Dataset #2 

```{r, include = F, echo = F}
data2 <- read_csv("Lab Data/Lab3/Ass5Data2.csv")
boxplot(QuailDens~Habitat, data = data2)
res2 <- lm(QuailDens~Habitat, data = data2)
summary(res2)
anova(res2)
confint(res2)
abline(res2)
resaov1 <- aov(QuailDens~Habitat, data = data2)
TukeyHSD(resaov1)
```
1. ANOVA indicated that at least two of the habitats were significantly different from each other (p = 0.0041).    

2. 

```{r, echo = FALSE, message = F}
boxplot(QuailDens~Habitat, data = data2)
```

3. We found that closed pine habitats had 0.025 (± 0.36; 95% C.I.) quail/ha less than ag habitats (p = 0.99) - *nonsignificant relationship*, while open pine habitats had 0.41 (± 0.36; 95% C.I.) quail/ha less than ag habitats (p = 0.024).  

## Dataset #3 

```{r, include = F, echo = F}
data3 <- read_csv("Lab Data/Lab3/Ass5Data3.csv")
boxplot(DeerKill~Hunt, data = data3)
res3 <- lm(DeerKill~Hunt, data = data3)
summary(res3)
anova(res3)
confint(res3)
abline(res3)
resaov2 <- aov(DeerKill~Hunt, data = data3)
TukeyHSD(resaov2)
```

1. ANOVA indicated that at least two of the habitats were significantly different from each other (p = 2.005e-10).  

2.  

```{r, echo = FALSE, message = F}
boxplot(DeerKill~Hunt, data = data3)
```

3. We found that hunters using a rifle had 1.91 (± 2.18; 95% C.I.) deer killed/hunter more than hunters using a bow (p = 0.094) *nonsignificant relationship*, while hunters using an Uzi had 8.57 (± 2.13; 95% C.I.) deer killer/hunter more than hunters using a bow (p = < 0.0001). Hunters using an Uzi had 6.66 (± 2.07; 95% C.I.) deer killer/hunter more than hunters using a rifle (p = < 0.0001).  

## Dataset #4 

```{r, include = F, echo = F}
data4 <- read_csv("Lab Data/Lab3/Ass5Data4.csv")
plot(Biomass~as.factor(Fertilizer), data = data4)
data4$Fertilizer <- factor(data4$Fertilizer, levels = c("2","0","4","6","8","10"))
res4 <- lm(Biomass~as.factor(Fertilizer), data = data4)
summary(res4)
```

```{r, include = F, echo = F}
data4cont <- read_csv("Lab Data/Lab3/Ass5Data4.csv")
plot(Biomass~Fertilizer, data = data4cont)
res4cont <- lm(Biomass~Fertilizer, data = data4cont)
summary(res4cont)
confint(res4cont)
data4$Fertilizer <- factor(data4$Fertilizer, levels = c("2","0","4","6","8","10"))
res4 <- lm(Biomass~as.factor(Fertilizer), data = data4)
anova(res4,res4cont)
```

1. We found that using 2 g fertilizer had 2.14 (± 1.50; 95% C.I.) kg/ha more biomass than plots with 0 g fertilizer in restored prairies (p = 0.0070).  
2. We found that using 4 g fertilizer had 2.18 (± 1.50; 95% C.I.) kg/ha mroe biomass than plots with 2 g fertilizer in restored prairies (p = 0.0061).  
3. We found that for each 1 g increase in fertilizer, we observed a 0.040 (± 0.25; 95% C.I.) kg/ha decrease in biomass (p = 0.75, r^2^ = 0.0038) *nonsignificant relationship*  
4. The f-drop F-statistic is 22.18 and p = 9.07e-08.  
5. *The model where Fertilizer is categorical is preferred.*  
6. The relationship between biomass and fertilizer is non-linear.  