---
title: "Lab 5"
author: "Steven Cabrera"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(readr)
library(ggplot2)
library(dplyr)
library(car)
library(carData)
library(emmeans)
```

## Dataset #1 

```{r, include = F, echo = F}
data1 <- read_csv("Lab Data/Lab5/Assignment4b.csv")
```
In this example, you have two continuous factors (e.g., elevation [meters] and degrees north latitude [degrees]), and one categorical factor with three groups (country). The dependent variable is lotus plant size (grams). In this system, there is no collinearity among your variables.

1. $Y = B_0 + B_1 RoadDensity + B_2 Habitat + B_3 RoadDensity * Habitat + E \sim \text{N}(\mu, \sigma)$  
    * X varibles:
        + $X_1$ = Latitude (continuous, degrees)
        + $X_2$ = Elevation (continuous, meters)
        + $X_3$ = Country (categorical, three groups = "India", "Nepal", "Tibet" ; reference = "India" in this case)
    * $B$ variables:
        + $B_0$: y-intercept
        + $B_1$: slope (estimate) of latitude
        + $B_2$: slope (estimate) of elevation
        + $B_3$: slope (estimate) of country
 