---
title: "Lab 8"
author: "Steven Cabrera"
date: "2022-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(readr)
library(ggplot2)
library(dplyr)
library(car)
library(carData)
library(nlme)
library(lme4)
```

```{r, include = F, echo = F}
data1 <- read_csv("Lab Data/Lab8/PeckerData2.csv")
res1 <- glm(data = data1, Present~Burned+StandAge+TreeDensity, family = binomial)
summary(res1)
exp(0.05555)
confint(res1)
```

1. We found that woodpecker detection was 4.78 (1.59 - 16.8; 95% C.L.) times as likely in burned stands compared to unburned stands (p = 0.008).  
For each 1 year increase in stand age, we observed a 5.7% (3.39 - 8.71%) increase in woodpecker detection (p = 1.09e-05).  
For each 1 tree/100m2 increase in tree density, we observed a 12% (5.9 - 18.8%) decrease in woodpecker detection (p = 0.0006).

```{r, include = F, echo = F}
data2 <- read_csv("Lab Data/Lab8/BirdData.csv")

x=seq(1,length(data2$Size),1)
```

2. For each 10 index increase in understory density, we observed a 11.0% (8.0 - 14.0%) increase in the number of birds found (p < 2e-16).    
We found that the number of birds found was 1.69 (1.41 - 2.03) times as likely in forested habitats compared to crop habitats (p = 1.88e-08).    
We found that the number of birds found was 1.78 (1.45 - 2.18) times as likely in grassland habitats compared to crop habitats (p =3.86e-08).    
Though non-signiicant, we found that the number of birds found was 1.05 (0.88 - 1.25) times as likely in grassland habitats compared to forested habitats (p = 0.583).  

```{r, include = F, echo = F}
data3 <- read_csv("Lab Data/Lab8/Conceptions.csv")
head(data3)
res3 <- glmer(data = data3, Conceived~Sperm+Frozen+(1|Individual), family = binomial)
summary(res3)
```

3. For each 1 million increase in sperm recieved, we observed a 1.04 (1.02 - 1.07) times as likely for rhinos to conceive  (p = 7.39e06).    
Though non-significant, we found that individuals were 0.56 (0.29 - 1.04) times as likely to be concieved if the sperm sample was cyro-frozen compared to fresh (p = 0.07).  

Standard deviation for the random effect was 0.3723.
